Figures for Paper

### 2021-02-11 paper_figures.Rmd
peptide plot and data

@@

fig1a = ggplot() +
  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions != 0),
             aes(x = deltaPSI, y =log10_test_stat,
                 alpha = graph_alpha,,fill = "Annotated Junction"), pch = 21, size = 4) + 
  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions == 0),
             aes(x = deltaPSI, y =log10_test_stat,
                 alpha = graph_alpha,fill = "Novel Junction"), pch = 21, size = 4) + 
  geom_text_repel(data = splicing_dots_tables[junction_name != ""],
                  aes(x = deltaPSI, y =log10_test_stat,
                      label = label_junction,
                      color = as.character(de_novo_junctions)), point.padding = 0.3,
                    nudge_y = 0.2,
                  min.segment.length = 0,
                  box.padding  = 2,
                  size=6,show.legend = F) +
  geom_hline(yintercept = -log10(1 - .9)) + 
     scale_fill_manual(name = "",
                      breaks = c("Annotated Junction", "Novel Junction"),
                      values = c("Annotated Junction" = "#648FFF", "Novel Junction" = "#fe6101") ) +
       scale_color_manual(name = "",
                      breaks = c("0", "1"),
                      values = c("1" = "#648FFF", "0" = "#fe6101") ) +
  guides(alpha = FALSE, size = FALSE) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab(expression(paste("-Lo", g[10], " Test Statistic"))) +
  theme(text = element_text(size = 24)) +
  theme(legend.text=element_text(size=22)) +
  xlim(-1,1)

@@

fig2a = ggplot(data = de_table) + 
  geom_point(data = de_table %>% filter(contains_cryptic == "0"),
         aes(x = log2FoldChange, y = -log10(padj),
             alpha = graph_alpha,fill = "No Cryptic"), pch = 21, size = 4) + 
    geom_point(data = de_table %>% filter(contains_cryptic == "1"),
         aes(x = log2FoldChange, y = -log10(padj),
             alpha = graph_alpha,fill = "Contains Cryptic"), pch = 21, size = 4) + 
    geom_text_repel(data = de_table[label_junction != ""],max.overlaps = 500,
                  aes(x = log2FoldChange, 
                      y = -log10(padj),
                      label = label_junction,
                      color = as.character(contains_cryptic)),
                   nudge_y = 5,
                  min.segment.length = 0,
                  box.padding  = 4,
                  size=6,show.legend = F) +
       scale_fill_manual(name = "",
                      breaks = c("No Cryptic", "Contains Cryptic"),
                      values = c("No Cryptic" = "#648FFF", "Contains Cryptic" = "#fe6101") ) +
         scale_color_manual(name = "",
                      breaks = c("1", "0"),
                      values = c("1" = "#fe6101", "0" = "#648FFF") ) +
  xlim(-7.5,7.5) + 
  ylab(expression(paste("-Lo", g[10], " P-value"))) +
    guides(alpha = FALSE, size = FALSE) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab(expression(paste("Lo", g[2], " Fold Change"))) +
  theme(text = element_text(size = 24)) +
  theme(legend.text=element_text(size=22)) +
    geom_hline(yintercept = -log10(0.1))  +
  geom_vline(xintercept=c(0), linetype="dotted")

@@

We have 2 PSI's for the UNC13A cryptic, one which includes both the short and
long form of the cryptic, and one which does not. While we're not mentioning in 
figure 1, given that the longer form of the cryptic appears in control cerebellum, 
one could argue that that junction should not be included in the PSI calculation
here for UNC13A cryptic per se. In practice, this makes very little difference in the
conclusions made on the cell lines, so I've included both for completeness.

The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192. 
The NB cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele. 
There was variability on the Klim hMN set on allelic expression. 
```{r}

rel_rna_cryptic_amount = data.table::fread(file.path(here::here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
####barplot UNC13A CE PSI - full five####
unc13a_cryptic_full = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A CE PSI") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

####barplot UNC13B NMD PSI####
unc13b_psi_nmd = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "unc13b_nmd_exon_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B \nNMD Exon PSI") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12) +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
####UNC13A RNA and UNC13A Cryptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12) +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) + 
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
      theme(text = element_text(size = 18,family = 'sans'), 
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
####UNC13A RNA and UNC13A Cryptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12) +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) + 
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
      theme(text = element_text(size = 18,family = 'sans'), 
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
####UNC13A RNA and UNC13A IR####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12) +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
####TARDBP RNA and UNC13A IR####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13a_ir),size = 12) +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom')

unca = rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "cryptic_psi_full",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("UNC13A CE PSI") + 
  xlab("") + 
  guides(color = FALSE)

stmn2 = rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "stmn_2_cryptic_psi",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("STMN2 Cryptic PSI") + 
  xlab("") + 
  guides(color = FALSE)

ggarrange(unc13a_cryptic_full,unc13b_psi_nmd,unc13a_ir,unc13b_ir,common.legend = T)
####scatterplot stmn2 normalized TARDBP####

rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = stmn_2_cryptic_psi, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = stmn_2_cryptic_psi)) +
    theme(legend.position = 'bottom') + 
  facet_wrap(~source)
####scatterplot UNC13BIR normalized TARDBP####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')
####scatterplot UNC13AIR and UNC13A Crptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13a_ir)) +
    theme(legend.position = 'bottom')
####scatterplot UNC13BIR and UNC13A Crptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)
####scatterplot UNC13BIR and UNC13B NMD####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)

####barplot UNC13A RNA level####

@@

# Paient summary statistics
First let's look at only including tissues with detectable stmn2 or UNC13A CE
```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(call) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi))

@@

(As a side note, once we do this the number of ALS-non-TDP drops down to 6 (2 FUS) because
the ALS sample tissues are not balanced and not every participant has samples in every tissue)
```{r}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_1",variable)) %>% 
    group_by(sample) %>% 
    mutate(inclusion_reads = sum(value)) %>% 
    ungroup() %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>% 
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```
# Detection of UNC13A
Looking at disease tissue only, so just taking the cords in ALS and the 
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.
```{r}

@@

# Detection UNC13A by tissue

>>>>

```{r volcano_plots}

@@

fig1a = ggplot() +
  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions != 0),
             aes(x = deltaPSI, y =log10_test_stat,
                 alpha = graph_alpha,,fill = "Annotated Junction"), pch = 21, size = 10) + 
  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions == 0),
             aes(x = deltaPSI, y =log10_test_stat,
                 alpha = graph_alpha,fill = "Novel Junction"), pch = 21, size = 10) + 
  geom_text_repel(data = splicing_dots_tables[junction_name != ""],
                  aes(x = deltaPSI, y =log10_test_stat,
                      label = label_junction,
                      color = as.character(de_novo_junctions)), point.padding = 0.3,
                    nudge_y = 0.2,
                  min.segment.length = 0,
                  box.padding  = 2,
                  size=6,show.legend = F) +
  geom_hline(yintercept = -log10(1 - .9)) + 
  geom_vline(xintercept = -0,linetype="dotted") + 
     scale_fill_manual(name = "",
                      breaks = c("Annotated Junction", "Novel Junction"),
                      values = c("Annotated Junction" = "#648FFF", "Novel Junction" = "#fe6101") ) +
       scale_color_manual(name = "",
                      breaks = c("0", "1"),
                      values = c("1" = "#648FFF", "0" = "#fe6101") ) +
  guides(alpha = FALSE, size = FALSE) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab(expression(paste("-Lo", g[10], " Test Statistic"))) +
  theme(text = element_text(size = 24)) +
  theme(legend.text=element_text(size=22)) +
  xlim(-1,1)

@@

fig1b = ggplot(data = de_table) + 
  geom_point(data = de_table %>% filter(contains_cryptic == "0"),
         aes(x = log2FoldChange, y = -log10(padj),
             alpha = graph_alpha,fill = "No Cryptic"), pch = 21, size = 10) + 
    geom_point(data = de_table %>% filter(contains_cryptic == "1"),
         aes(x = log2FoldChange, y = -log10(padj),
             alpha = graph_alpha,fill = "Contains Cryptic"), pch = 21, size = 10)+ 
    geom_text_repel(data = de_table[label_junction != ""],max.overlaps = 500,
                  aes(x = log2FoldChange, 
                      y = -log10(padj),
                      label = label_junction,
                      color = as.character(contains_cryptic)),
                   nudge_y = 5,
                  min.segment.length = 0,
                  box.padding  = 4,
                  size=6,show.legend = F) +
       scale_fill_manual(name = "",
                      breaks = c("No Cryptic", "Contains Cryptic"),
                      values = c("No Cryptic" = "#648FFF", "Contains Cryptic" = "#fe6101") ) +
         scale_color_manual(name = "",
                      breaks = c("1", "0"),
                      values = c("1" = "#fe6101", "0" = "#648FFF") ) +
  xlim(-7.5,7.5) + 
  ylab(expression(paste("-Lo", g[10], " P-value"))) +
    guides(alpha = FALSE, size = FALSE) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab(expression(paste("Lo", g[2], " Fold Change"))) +
  theme(text = element_text(size = 24)) +
  theme(legend.text=element_text(size=22)) +
    geom_hline(yintercept = -log10(0.1))  +
  geom_vline(xintercept=c(0), linetype="dotted")

print(fig1a)
print(fig1b)
```

The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192. 
The SK-N-DZ cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele. 
There was variability on the Klim hMN set on allelic expression.

```{r read in }

rel_rna_cryptic_amount = data.table::fread(file.path(here::here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
```

## barplot UNC13A CE PSI - full five

```{r}
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A CE PSI") + 
    xlab("") + 
    guides(color = FALSE) +
    scale_y_continuous(labels = scales::percent) +
    theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))
```

## barplot UNC13B NMD PSI

```{r}
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "unc13b_nmd_exon_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B \nNMD Exon PSI") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))
```

#### TARDBP RNA and UNC13A Cryptic
```{r}

rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12) +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
```

#### UNC13A RNA and UNC13A Cryptic####

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12,
             method = 'spearman',
           cor.coef.name = 'rho') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
      scale_y_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) + 
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
      theme(text = element_text(size = 18,family = 'sans'), 
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
```
#### UNC13A RNA and UNC13A IR
```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12,cor.coef.name = 'rho',method = 's') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
```
#### TARDBP RNA and UNC13A IR

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13a_ir),size = 12) +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
```

#### barplot UNC13A relative RNA across experiments

```{r}
rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "cryptic_psi_full",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("UNC13A CE PSI") + 
  xlab("") + 
  guides(color = FALSE) 
```

```{r}
stmn2 = rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "stmn_2_cryptic_psi",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("STMN2 Cryptic PSI") + 
  xlab("") + 
  guides(color = FALSE)

####scatterplot stmn2 normalized TARDBP####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = stmn_2_cryptic_psi, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = stmn_2_cryptic_psi)) +
    theme(legend.position = 'bottom') 
####scatterplot UNC13BIR normalized TARDBP####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')
####scatterplot UNC13AIR and UNC13A Crptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13a_ir)) +
    theme(legend.position = 'bottom')
####scatterplot UNC13BIR and UNC13A Crptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)
```

#### scatterplot UNC13BIR and UNC13B NMD

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)

@@

#### barplot protein abundances iPSC KD

```{r}
peptides = fread(file.path(here::here(),"data","peptide_abdundance.csv"))

peptides %>% 
  mutate(gene = fct_relevel(gene, "UNC13A","UNC13B")) %>% 
  ggbarplot(,
            x = "condition",
            add = c("mean_se"),
            y = "protein",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8),
            facet.by = 'gene') + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("Protein abundance") + 
  xlab("") + 
  geom_jitter(pch = 21,height = 0,aes(fill = condition),size = 2) + 
  guides(color = FALSE) +
  facet_wrap(~gene, scales = 'free') +
  theme(legend.position = 'none') + 
  scale_y_continuous(labels = scientific) + 
  stat_compare_means(comparisons = list(c("Control","TDP-43 KD")),
                     label = 'p.signif',tip.length = 0,
                     size = 10)

@@

# Patient summary statistics

```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(call) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi))

@@

(As a side note, once we do this the number of ALS-non-TDP drops down to 6 (2 FUS) because
the ALS sample tissues are not balanced and not every participant has samples in every tissue)

```{r}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_1",variable)) %>% 
    group_by(sample) %>% 
    mutate(inclusion_reads = sum(value)) %>% 
    ungroup() %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>% 
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```
# Detection of UNC13A
Looking at disease tissue only, so just taking the cord and motor cortex in ALS and the 
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.

@@

# UNC13A detection rate by tissue

```{r}
boxplot_table %>% 
  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(sample,disease_group2,detected,tissue_clean) %>% 
  unique() %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  ungroup() %>% 
  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>% 
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Motor Cortex",
                                    "Thoracic Spinal Cord")) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate,fill = detection_name),position = position_dodge2(width = 0.8, preserve = "single")) + 
  ggpubr::theme_pubr() + 
  ylab("Percent of Tissues \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N tissues") + 
  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) + 
  theme(axis.text.x =  element_text(size = 14)) + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  +
  scale_fill_manual(values = c("#4C97B5","#4C97B5","#4C97B5","#4C97B5","#4C97B5","red","#C87156","#C87156")) +
  theme(legend.position = "none") 
```
